---
title: "blblm Package"
author: "Diana Rueda"
date: "03/17/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{blblm_project_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(blblm)
```

Implements linear regression analysis through Bag of Little Bootstraps procedure which combines features from bootstrap and subsampling. This package has the option to use multiple cores for faster processing.     
Bag of little bootstrap consists on sampling n items from b items available with replacement. It samples s times into sizes of b. Then for each subsample it resamples each r times until sample size is n. It then computes the bootstrap statistic for each bootstrap sample and the statistic from the bootstrap statistics. Finally it takes the average of the statistics.

---------------------------

# Functions

### blblm
Builds blblm object      
      
Parameters:       
**formula:** Linear regression formula of the form y ~ x. It can be multivariable linear regression analysis.   
**data:** Dataframe containing response and independent variables y and x.      
**m:** Integer indicating the number of items to choose from.     
**B:** Integer indicating the number of times to replicate the linear regression fitting process.    
**workers:** Integer indicating the number of workers. If workers has a value larger than one the process will run in paralelization.       
**fastlm:** If true the algorithm will run with function fastLm if false it will use lm.wfit.    
    
Returns: object class blblm     
     
Creates linear regression objects with little bag of bootstraps. Splits the data into subsamples, and uses future_map when paralization is chosen and map when only using one core to apply the indicated lm function to the subsample. 


### split_data

Splits data into m parts of approximated equal sizes. Called through blblm() for bootstrap sampling.      
      
Parameters:       
  
**data:** Dataframe containing response and independent variables y and x.      
**m:** Integer indicating the number of items to choose from.     
    
Returns: List of data frames representing subsamples  
     
Uses ```sample.int(m, nrow(data), replace = TRUE)``` to reorganize the data given into subsamples. Then uses ```split()``` to split the observations into groups.    


### lm_each_subsample
Computes the regression model estimates, and replicates lm1 process. Called through ```blblm()``` as mapping funciton.      
      
Parameters:       
**formula:** Linear regression formula of the form y ~ x. It can be multivariable linear regression analysis.   
**data:** Dataframe containing response and independent variables y and x.      
**m:** Integer indicating the number of items to choose from.     
**B:** Integer indicating the number of times to replicate the linear regression fitting process.    
**workers:** Integer indicating the number of workers. If workers has a value larger than one the process will run in paralelization.       
**fastlm:** If true the algorithm will run with function fastLm if false it will use lm.wfit.    
    
Returns: ‘pdb’ with replicated atomic coordinates   
     
Strctures regression model with methods ```.frame()```, ```.matrix()``` and ```.response()``` for each subsample and replicates lm1 ```B``` times without simplification.    


### lm1
Computes the regression estimates for a blb dataset. Called through lm_each_subsample.           
      
Parameters:       
   
**X:** Data with independent variables obtained from ```model.matrix()```    
**y:** Vector of response data obtained from ```model.response()```       
**n:** Number of rows of complete data before subsampling.       
**fastlm:** If true the algorithm will run with function fastLm if false it will use lm.wfit.    
    
Returns: list of model coefficients  
     
Calculates weigths to simulate distribution with the help of ```rmultinom()```, and fits model either through lm.wfit or fastLm.
Then, calls ```blbcoef()``` and ```sigma()``` to compute fit coefficients and places them in a list.



### blbcoef
Computes the coefficients from model fit           
      
Parameters:       
   
**fit:** Fitted blblm model     
    
Returns: list of model coefficients  
     
Obtains blblm model coefficients through ```coef()```.     


### blbsigma
Computes sigma from model fit           
      
Parameters:       
   
**fit:** Fitted blblm model     
    
Returns: numeric sigma value 
     
Calculates sigma from formula with use of model rank, residuals and weights.        


----------------------------

# Methods

### print.blblm

```
print.blblm <- function(x, ...) {
  cat("blblm model:", capture.output(x$formula))
  cat("\n")
}
```


### sigma.blblm

Calculates sigma if ```confidence = FALSE ``` and returns sigma confidence interval if ```confidence = TRUE```.    

```
sigma.blblm <- function(object, confidence = FALSE, level = 0.95, ...) {
  est <- object$estimates
  sigma <- mean(map_dbl(est, ~ mean(map_dbl(., "sigma"))))
  if (confidence) {
    alpha <- 1 - 0.95
    limits <- est %>%
      map_mean(~ quantile(map_dbl(., "sigma"), c(alpha / 2, 1 - alpha / 2))) %>%
      set_names(NULL)
    return(c(sigma = sigma, lwr = limits[1], upr = limits[2]))
  } else {
    return(sigma)
  }
}
```

### coef.blblm

Calculates coefficients estimates through row means and reduce.

```
#' @export
#' @method coef blblm
coef.blblm <- function(object, ...) {
  est <- object$estimates
  map_mean(est, ~ map_cbind(., "coef") %>% rowMeans())
}
```

### confint.blblm

Calculates confidence interval for coefficient estimates.


```
confint.blblm <- function(object, parm = NULL, level = 0.95, ...) {
  if (is.null(parm)) {
    parm <- attr(terms(object$formula), "term.labels")
  }
  alpha <- 1 - level
  est <- object$estimates
  out <- map_rbind(parm, function(p) {
    map_mean(est, ~ map_dbl(., list("coef", p)) %>% quantile(c(alpha / 2, 1 - alpha / 2)))
  })
  if (is.vector(out)) {
    out <- as.matrix(t(out))
  }
  dimnames(out)[[1]] <- parm
  out
}
```

### predict.blblm

Calculates prediction from new data. If ```confidence = TRUE``` prediction interval is returned.

```
predict.blblm <- function(object, new_data, confidence = FALSE, level = 0.95, ...) {
  est <- object$estimates
  X <- model.matrix(reformulate(attr(terms(object$formula), "term.labels")), new_data)
  if (confidence) {
    map_mean(est, ~ map_cbind(., ~ X %*% .$coef) %>%
      apply(1, mean_lwr_upr, level = level) %>%
      t())
  } else {
    map_mean(est, ~ map_cbind(., ~ X %*% .$coef) %>% rowMeans())
  }
}
```

----------------------------

# Results

```{r}
library(blblm)
library(future)
library(furrr)
library(ggplot2)
library(bench)
```

The following code generates data to use in blblm to benchmark time efficiency.    

```{r}
library(tibble)
y = rnorm(100000, mean = 15, sd = 3)
x1 = 0.4 * y + rnorm(100000, sd = 0.25)
x2 = 0.7 * y + rnorm(100000, sd = 0.25)
X = tibble(x1,x2,y)

```

The first result saved in blblm is the result of using the configuraiton of the blblm package before any alterations for efficiency. The result saved in blblm_4wkrs is the result of the configuration where we run the blb process with 4 workers and fastLm. Bellow are the results.

```{r paged.print = FALSE}
result = bench::mark(
  blblm = blblm(y ~ x2 * x1, data = X, m = 5, B = 1000, workers = 1, fastlm = FALSE),
  blblm_4wkrs = blblm(y ~ x2 * x1, data = X, m = 5, B = 1000, workers = 4, fastlm = TRUE),
  relative = FALSE,
  check = FALSE,
  iterations = 10
)
result
```



----------------------------

# Example

```{r}
fit <- blblm(mpg ~ wt * hp, data = mtcars, m = 3, B = 100)
coef(fit)
```

```{r}
confint(fit, c("wt", "hp"))
```

```{r}
sigma(fit)
```

```{r}
sigma(fit, confidence = TRUE)
```

```{r}
predict(fit, data.frame(wt = c(2.5, 3), hp = c(150, 170)))
```

```{r}
predict(fit, data.frame(wt = c(2.5, 3), hp = c(150, 170)), confidence = TRUE)

```




